Linear Algebra as a Unifying Theory\\of Wireless Workload Modeling

### 2008-02-04 paper.tex
Switching over to coding...

The traffic workload pattern in a network is simply a collection of \caps{IP} packets between hosts in the network sent at certain times. It is standard in network analysis, however, to aggregate sequences of packets sharing the same \caps{IP} protocol, source and destination nodes, and source and destination \caps{TCP} or \caps{UDP} port numbers. Such a sequence of packets is called a \textit{flow}. The traffic pattern is characterized completely by the detailed behavior of it's flows. The behavior of a flow is defined by six properties: its source and destination nodes; its start and end times; and the specific sizes of and intervals between its constituent packets. We propose to represent the behavior of each flow in terms of six vectors---one for each aspect of its behavior. The aggregate behavior of all the flows in a workload pattern are then represented as matrices, the rows of which are the behavior vectors of the flows.

@@

It is common in generating traffic behaviors for experiments to treat packets, nodes, and flows uniformly to varying degrees, largely because of the high complexity of doing otherwise and the lack of compelling alternatives. Here we examine how various notions of uniformity are expressed in the matrix model. In general, \textit{intra-flow uniformity} is effected by right multiplication, which mixes the elements in each row, whereas \textit{inter-flow uniformity} is effected by left multiplication, which mixes elements across rows. The constant bit-rate (\caps{CBR}) flow model, for example, in its strictest sense, is an intra-flow model, which stipulates that all the packets in each flow behave identically, having identical sizes and intervals; thus each flow's bit-rate is constant, as the name implies. The strict \caps{CBR} model can be expressed simply using the packets-bytes-duration (\caps{PBD}) model used as an example in \Section{representations}. The pair of matrices $\seq{Z,V}$ is transformed to \caps{PBD} form using the vectors defined in \Section{flow-totals}: $\seq{Z,V}\Gamma$, where
\begin{align}
\Gamma=\begin{pmatrix}
\ones{} & \vec{\sigma} & \zeros{} \\
\zeros{} & \zeros{} & \vec{\tau}
\end{pmatrix} \in \R^{(d_v+d_z)\times{3}}.
\end{align}

@@

While constant bit-rate in the strict sense implies only that the packet behavior in each flow is uniform, it is extremely rare for it to be applied in this strict sense only; it is typical that \caps{CBR} models make assumptions about uniformity across flows, and not only within each flow. The most drastic simplification is to treat all flows in a synthetic workload identically. This is effected by left multiplication: $F'=\ones{f \times f} F$. In general, we can express uniformity through left-multiplication by some matrix. To make flow behavior uniform at the per-source node level, we left-multiply by $SS^t/f\in\R^{f \times f}$: $F'=(SS^t/f)F$. Per-destination node uniformity can be achieved similarly: $F'=(DD^t/f)F$.

Although assumptions about uniformity across flows are common, there is little agreement about which notion of uniformity is most appropriate; often it is simply left unspecified. The confusion arises due to the lack of a unique way of associating flows with nodes. Each node can equally viably be associated with its source or its destination. Which association is more valid? Consider a web request: should the behavior of the request be determined by what the client typically does? Or should the requests that the server typically receives determine behavior? There is no armchair answer: only analysis of real data can give us the answer.

>>>>

The traffic workload pattern in a network is a collection of \caps{IP} packets between hosts in the network sent at certain times. It is standard in network analysis, however, to aggregate sequences of packets sharing the same \caps{IP} protocol, source and destination nodes, and source and destination \caps{TCP} or \caps{UDP} port numbers. Such a sequence of packets is called a \textit{flow}. The traffic pattern is characterized completely by the detailed behavior of it's flows. The behavior of a flow is defined by six properties: its source and destination nodes; its start and end times; and the specific sizes of and intervals between its constituent packets. We propose to represent the behavior of each flow in terms of six vectors---one for each aspect of its behavior. The aggregate behavior of all the flows in a workload pattern are then represented as matrices, the rows of which are the behavior vectors of the flows.

@@

It is common in generating traffic behaviors for experiments to treat packets, nodes, and flows uniformly to varying degrees, largely because of the high complexity of doing otherwise and the lack of compelling alternatives. Here we examine how various notions of uniformity are expressed in the matrix model. In general, \textit{intra-flow uniformity} is effected by right multiplication, which mixes the elements in each row, whereas \textit{inter-flow uniformity} is effected by left multiplication, which mixes elements across rows. The constant bit-rate (\caps{CBR}) flow model, for example, in its strictest sense, is an intra-flow model, which stipulates that all the packets in each flow behave identically, having identical sizes and intervals; thus each flow's bit-rate is constant, as the name implies. The strict \caps{CBR} model can be easily expressed with the packets-bytes-duration (\caps{PBD}) model used as an example in \Section{representations}. The pair of matrices $\seq{Z,V}$ is transformed to \caps{PBD} form using the vectors defined in \Section{flow-totals}: $\seq{Z,V}\Gamma\in\R^{f \times 3}$, where
\begin{align}
\Gamma=\begin{pmatrix}
\ones{} & \vec{\sigma} & \zeros{} \\
\zeros{} & \zeros{} & \vec{\tau}
\end{pmatrix} \in \R^{(d_v+d_z)\times{3}}.
\end{align}

@@

While constant bit-rate in the strict sense implies only that the packet behavior in each flow is uniform, it is very rare for it to be applied in this strict sense only; it is typical that ``\caps{CBR}'' models make assumptions about uniformity across flows, and not just within each flow. The most drastic simplification is to treat all flows in a synthetic workload identically. This is effected by left multiplication: $F'=(\ones{f \times f}/f)F$. To make flow behavior uniform at the per-source node level, we left-multiply by $SS^t/f\in\R^{f \times f}$: $F'=(SS^t/f)F$. Per-destination node uniformity can be achieved similarly: $F'=(DD^t/f)F$.

Although assumptions about uniformity across flows are common, there is little agreement about which notion of uniformity is most appropriate; often it is simply left unspecified. The confusion arises due to the lack of a unique way of associating flows with nodes. Each node can equally viably be associated with its source or its destination. Which association is more valid? Consider a web request: should the behavior of the request be determined by what the client typically does? Or should the requests that the server typically receives determine behavior? There is no armchair answer: only analysis of real data can give us the answer. The matrix model provides an analytical framework for doing so.
